---
title: "CA - S3 - Solutions"
Author: Josep Curto, IE Business School 
output:
  html_notebook:
    toc: yes
    toc_float: yes
---

# Warm-up Exercise

- Can you explain the meaning of the following code?

```{r}
1==1
0==1

testVar <- -5
testVar

testVar2 <- 9
testVar2

ifelse(abs(testVar)>3,print("YESSSS"),print("NOOO"))

if(testVar == 1 | testVar2 > 10)
{
  print("I am inside the if")
}else if(testVar==3)
{
  print("I am printing the else because I have the value three")
} else
{
  print("there is no hope...")
}
```

# Exercises (slide 10)

- Write a repeat{} loop that prints all the even numbers from 2 to 10. Start with i <- 0

```{r}
i <- 0

repeat{
  i <- i + 2
  print(i)
  if(i == 10) {
    break
  }
}
```

- Write a while() loop that increments the	variable, “i“, 6 times, and prints “msg” at 	every iteration using the following variables: 
    - msg <- c("Hello IE")
    - i <- 1

```{r}
msg <- c("Hello IE")
i <- 1

while (i < 7) {
  print(msg)
  i = i + 1
}
```

- Write a for() loop that prints the first four numbers of this sequence: x <- c(7, 4, 3, 8, 9, 25)

```{r}
x <- c(7, 4, 3, 8, 9, 25)

for (i in x) {
  if(i%%2==0)
  {
    print(i)
  }
}
```

# Exercises (slide 12)

- Create a function that will return the sum of 2 numbers

```{r}
FunctSum<- function (x, y) {
  result <- x + y
  return(result)
}

x<-FunctSum(5.5,8)
x
```

- Create a function what will return TRUE if a given integer is inside a vector
    - With while
    - With for
    - With %in%
    
```{r}
# With While
f.exists <- function (vec, value) {
  flag.exist <- FALSE
  i <- 1
  
  while (i <= length (vec) & !flag.exist) {
    
    if (vec[i] == value) {
      flag.exist <- TRUE
    }
    i <- 1 + i
  }
  return(flag.exist)
}
f.exists(c(1:10), 10)

# With For
for.exists<-function(vec,value)
{
  flag.exists=FALSE
  for (i in vec)
  {
    if (i == value)
    {
      flag.exists=TRUE
    }
  }
  return(flag.exists)
}

for.exists(c(1:10), "6")

# With %in%
check.wrap<-function(vec,x)
{
  flag.exists<- x %in% vec
 return (flag.exists)
}

check.wrap(c(1:10), "6")

```

- Create a function that given a data frame will print by screen the name of the column and the class of data it contains (e.g. Variable1 is Numeric).

```{r}
f.class <- function (df) {
  for (i in 1:ncol(df)) {
    print(paste(names(df)[i], "is", class(df[, i]),sep=" ") )
  }
}
f.class(cars)
f.class(iris)
```

- Create a function that given an integer will calculate how many divisors it has (other than 1 and itself). Make the divisors appear by screen.

```{r}
f.div <- function(n) {
  i <- 2
  counter <- 0
  
  while(i <= n/2) {
    if(n%%i==0) {
      counter <- counter + 1
      print(sprintf("%s is a divisor of %s",i,n))
    }
    i <- i + 1
  }
  sprintf("%s has %s divisors (other than 1 and itself)",n,counter)
}
f.div(16)
f.div(27)
f.div(53)
```

# Bonus Exercises

- Write a function that performs the union of two sets. Note: for the sake of simplicity, let's consider a vector as a set.

```{r}
set.union <- function(a, b) {
  u <- a
  for (i in 1:length(b)) {
    if (!(b[i] %in% u)) {
      u <- append(u, b[i])
    }
  }
  return(u)
}

a <- c(3, 5, 7, 11)
b <- c(3, 5, 13, 17)
set.union(a, b)
```

- Write a function that performs the intersection of two sets

```{r}
set.intersection <- function(a, b) {
  intersect <- vector()
  
  for (i in 1:length(a)) {
    if (a[i] %in% b) {
      intersect <- append(intersect, a[i])
    }
  }
  return(intersect)
}

a <- c(3, 5, 7, 11, 13, 20, 30)
b <- c(3, 5, 13, 17, 7, 10)

set.intersection(a, b)
```

- Create a function that tests if a set is a subset of another set.

```{r}
issubset <- function(a, b) {
  for (i in 1:length(a)) {
    if (!(a[i] %in% b)) {
      return(FALSE)
    }
  }
  return(TRUE)
}

a <- c(3, 5, 7, 11)
b <- c(3, 5, 13, 17)

c <- set.union(a, b)
c

print(issubset(a, c))
print(issubset(b, c))
print(issubset(c(3, 5, 7, 4), a))
```

# Bonus Exercises (II)

- Load s3.cvs

```{r}
customerData <- read.csv("s3.csv")
customerData 
```

- Create plot that represents the number of transactions per customer using ggplot2

```{r}
library(ggplot2)
ggplot(customerData, aes(factor(customerID)))+ geom_bar() +
  labs(x="Customer",y="# Transactions", title = "Transactions per Customer")
```

We have so many transactions and customers that the graph is really hard to read. Let's create a table with the aggregated number of transactions per customers.

```{r}
library(dplyr)
countTransactions <- count(customerData, customerID)
countTransactions
```


