---
title: "CA - S2 - Solutions"
output:
  html_notebook:
    theme: default
    toc: yes
    toc_float: yes
---

# Class Exercises 1

- Create a variable and store the number 25

```{r}
test <- 25
test
```

- Calculate the square root of 25

```{r}
test <- sqrt(test)
test
```

Another option:

```{r}
test <- sqrt(25)
test
```

- Create a vector with the first 100 numbers

```{r}
first100 <- 1:100
first100
```

We have other options such as:

```{r}
first100 <- seq(1:100)
first100 <- c(1:100)
```


- Create a vector to hold the data of a student (grade, course and class, all of them numeric)
    + access the element with the name “grade”
    + change the data to character data

```{r}
student <- c(grade=45,course=10,class=34)
student["grade"]
student[2] <- as.character(student[2])
class(student)
```

- Create a vector with the first 50 numbers
```{r}
first50 <- 1:50
first50 
```
- Create another vector withe the numbers 51-100
```{r}
second50 <- 51:100
second50
```
- Add them together
```{r}
vectorsum <- first50+second50
vectorsum
```

- Multiply them
```{r}
vector.product <- first50*second50
vector.product
```
- Join both of them in a larger vector
```{r}
total.vector <- c(first50,second50)
total.vector 
```

# Class Exercises 2

- Load the test dataset. Use: load("dataTB.rdatq")
```{r}
load("dataTB.rdatq")
```

- What is the median of the final price?
```{r}
median(dataTickets$final_price)
```

- What is the relationship between final price and user price?
```{r}
print(dataTickets$final_price/dataTickets$usr_price)
cor(dataTickets$usr_price,dataTickets$final_price)
plot(dataTickets$usr_price,dataTickets$final_price)
```

- What is the event with the highest final price?
```{r}
print(max(dataTickets$final_price))
dataTickets[dataTickets$final_price==max(dataTickets$final_price),]
```

# Homework Exercises

- Create a vector of coefficients, called coeffs, using the sample function. Here, we draw a sample of size 3 from −20, −19, . . . , 19, 20 with replacement
```{r}
coeffs <- sample(-20:20,3,replace=T)
coeffs 
```

- Determine the class of the object coeffs.
```{r}
class(coeffs)
```

- Determine the length of the object coeffs.
```{r}
length(coeffs)
```

- Determine the names associated with the vector
```{r}
names(coeffs)
```
- Assign some names (a,b,c) to the vector. 

```{r}
names(coeffs) <- c("a","b","c")
```

-  What is the class of the names of the vector coeffs?
```{r}
class(names(coeffs))
```