---
title: "CA - S5: Example"
author: Josep Curto, IE Business School
abstract: "This document introduces how to calculate RFM with R and how to create and understand customer segments based on RFM."
keywords: "r, rfm"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output: 
  html_notebook: 
    fig_caption: yes
    toc: yes
    toc_float: yes
    self_contained: yes
---

# Calculate RFM with R

## Load packages

We will use readr, dplyr and ggplot2 packages.

```{r packages, warning=FALSE, echo=FALSE, message=FALSE}
# List of packages for session
.packages = c("readr", "dplyr","ggplot2")

# Install CRAN packages (if not already installed)
.inst <- .packages %in% installed.packages()
if(length(.packages[!.inst]) > 0) install.packages(.packages[!.inst])

# Load packages into session 
lapply(.packages, suppressPackageStartupMessages(require), character.only=TRUE)
```

## Load functions

We are loading a function that calculates RFM Score.

```{r functions}
source("RFM.R")
```

## Loading data

We load data into a dataframe:

```{r loading, warning=FALSE, echo=FALSE, message=FALSE}
df <- read_csv("data/s5.csv")
df
```

We need to understand the structure of the data set:

```{r structure}
str(df)
```

We can review the main statistics:

```{r summary}
summary(df)
```

**Question: What we can say about the summary?**

# RFM

##  Calculate RFM

```{r RFM}
dfRFM <- rfm(df,
           analysisDate = "2015/01/01", 
           customer = "customerID", 
           date = "date", 
           revenue = "amount")
dfRFM
```

How many customers do we have in every group?

```{r countRFM}
countRFM <- count(dfRFM, RFM_Score)
countRFM
```

**Question: what we can observe?**

## Visual Analysis

```{r plot1}
ggplot(dfRFM, aes(factor(RFM_Score))) +
  geom_bar() +
  ggtitle('Customer Distribution per RFM') +
  labs(x="RFM",y="# Customers") + 
  theme(plot.title = element_text(color="#666666", face="bold", size=16, hjust=0)) +
  theme(axis.title = element_text(color="#666666", face="bold")) +
  theme(axis.text.x  = element_text(angle=90,vjust=0.5))
```

**Question: what we can observe?**

# Segmentation

## Based on Recency

Let's use **Recency** to segment our customers:

```{r segmentation}
dfRFM$segment <- NA
dfRFM$segment[which(dfRFM$Recency > 365/2)] = "Inactive"
dfRFM$segment[which(dfRFM$Recency <= 365/2 & dfRFM$Recency > 365/4)] = "Sleeping"
dfRFM$segment[which(dfRFM$Recency <= 365/4)] = "Active"
dfRFM
```

Let's review the outcome:
```{r customerSegmentation}
customerSegmentation <- count(dfRFM, segment)
customerSegmentation
```

Let's create a graph for each one of the groups:

```{r plot2}
ggplot(dfRFM, aes(factor(RFM_Score),fill =factor(segment))) +
  geom_bar() +
  labs(x="RFM",y="# Customers", title="Customer Distribution per Recency") + 
  theme(axis.text.x  = element_text(angle=90,vjust=0.5,size=8)) + 
  facet_grid(. ~ segment) + theme(legend.position="none")
```

Let's create another graph where we combine all the groups:

```{r plot 3}
ggplot(dfRFM, aes(factor(RFM_Score),fill =factor(segment))) +
  geom_bar() +
  labs(x="RFM",y="# Customers",title="Customer Distribution per Recency") +
  theme(axis.text.x  = element_text(angle=90,vjust=0.5,size=8)) + 
  theme(legend.title=element_blank())
```

**Question: what we can observe?**

# Saving the results

```{r write}
write.csv(dfRFM,"data/RFM.csv")
```

# Exercises

Q1. Can you create a segmentation for frequency and monetary?
Q2. Can you create a segmentation considering all RFM values? Hint: use the segments presented in the article.